High-Dimensional Regularized Discriminant Analysis

### 2014-01-01 reports/paper-hdrda.Rnw
More polish applied to the summary of the model selection.

      Compute $\#\{\text{Error}_v(\lambda, \gamma)\} \leftarrow$ misclassified test observations
    }   
   }
  Compute $\widehat{\text{Error}}(\lambda, \gamma) = N^{-1} \sum_{v=1}^V \#\{\text{Error}_v(\lambda, \gamma)\}$

@@

A primary contributing factor to the efficiency of Algorithm
\ref{alg:hdrda_model_selection} is our usage of the compact singular value
decomposition (SVD) (sometimes called the reduced or fast SVD). Rather than
computing the eigenvalue decomposition of $\widehat{\bm \Sigma}$ to obtain $\bm
U_1$, we instead obtain $\bm U_1$ by computing the eigendecomposition of a much
smaller $N \times N$ matrix when $p \gg N$
\cite[Chapter~18.3.5]{Hastie:2008dt}. Recall that the pooled sample covariance
matrix is $\widehat{\bm \Sigma} = \bm X^{T} \bm X$\footnote{TODO: $1/N$ is
  missing. Add this. It will be absorbed into the eigenvalues.}  Then, because
$\bm X = \bm M \bm D^{1/2} \bm U^{T}$ by the SVD, we have the eigendecomposition
$\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$, where $\bm M \in \mathbb{R}_{N
  \times N}$ is orthogonal, $\bm D^{1/2} \in \mathbb{R}_{N \times N}^{\ge}$ is a
diagonal matrix consisting of the singular values of $\bm X$, and $\bm U \in
\mathbb{R}_{p \times N}$ is orthogonal. Now, $\bm X \bm X^{T} = \bm M \bm D \bm
M^{T}$ is the eigenvalue decomposition of $\bm X \bm X^{T}$, where $\bm D$ is
the matrix of eigenvalues of $\bm X \bm X^{T}$ and the columns of $\bm M$ are
the corresponding eigenvectors. Hence, we obtain $\bm M$ and $\bm D$ from an
eigendecomposition of $\bm X \bm X^{T} \in \mathbb{R}_{N \times N}$. Next, we
compute $\bm U = \bm X^{T} \bm M \bm D^{+/2}$, where $\bm D^{+/2} = \bm
D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the number of numerically
nonzero eigenvalues present in $\bm D$, by calculating the number of eigenvalues
that exceeds some tolerance value, say, $\epsilon = 1 \times 10^{-6}$. After
calculating $q$, we extract $\bm U_1$ by retaining only the first $q$ columns of
$\bm U$.

As a result of the compact SVD, we need calculate $\bm X_c \bm U_1$ only once
per cross-validation fold, requiring $O(p q N) \approx (p N^2)$ calculations.
Hence, the computational costs of expensive calculations, such as matrix
inverses and determinants, are greatly reduced because they are performed in the
$q$-dimensional subspace. Similarly, we reduce the dimension of the test data
set by calculating $\bm X_{test} \bm U_1$ once per fold. Conveniently, we see
that the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is
$\bm X_k \bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after
the initial calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm
Q_k$ requires $O(TODO)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times
  n_k}$, both its determinant and inverse require $O(n_k^3)$
operations. Consequently, $\bm W_k^{-1}$ requires $O(TODO)$ operations. Also,
the inverse and determinant of the diagonal matrix $\bm \Gamma_k \in
\mathbb{R}_{q \times q}$ requires $O(q)$ operations. Therefore, the $| \bm W_k
|$ requires $O(q n_k^3)$ operations.

The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
the selection of $\lambda$ and $\gamma$ via cross-validation because the most
time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
and $\gamma$.

>>>>

      Compute the number of misclassified test observations $\#\{\text{Error}_v(\lambda, \gamma)\}$ 
    }   
   }
  Compute $\widehat{\text{Error}}(\lambda, \gamma) = N^{-1} \sum_{v=1}^V \#\{\text{Error}_v(\lambda, \gamma)\}$

@@

A primary contributing factor to the efficiency of Algorithm
\ref{alg:hdrda_model_selection} is our usage of the compact singular value
decomposition (SVD) (sometimes called the reduced or fast SVD). Rather than
computing the eigenvalue decomposition of $\widehat{\bm \Sigma}$ to obtain $\bm
U_1$, we instead obtain $\bm U_1$ by computing the eigendecomposition of a much
smaller $N \times N$ matrix when $p \gg N$
\cite[Chapter~18.3.5]{Hastie:2008dt}. Applying the SVD, we decompose $\bm X_c =
\bm M \bm \Delta \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times p}$ is
orthogonal, $\bm \Delta \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix
consisting of the singular values of $\bm X_c$, and $\bm U \in \mathbb{R}_{p
  \times p}$ is orthogonal. Recalling that $\widehat{\bm \Sigma} = N^{-1} \bm
X_c^{T} \bm X_c$, we have the eigendecomposition $\widehat{\bm \Sigma} = \bm U
\bm D \bm U^{T}$, where $\bm U$ is the matrix of eigenvectors of $\widehat{\bm
  \Sigma}$ and $\bm D = N^{-1} \bm \Delta$ is the diagonal matrix of eigenvalues
of $\widehat{\bm \Sigma}$. Now, we can obtain $\bm M$ and $\bm D$ efficiently
from the eigenvalue decomposition of $\bm X_c \bm X_c^{T} = \bm M \bm D \bm
M^{T}$. Hence, we eigendecomposition of $\bm X_c \bm X_c^{T} \in \mathbb{R}_{N
  \times N}$. Next, we compute $\bm U = \bm X_c^{T} \bm M \bm D^{+/2}$, where
$\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the
number of numerically nonzero eigenvalues present in $\bm D$, by calculating the
number of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1
\times 10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only
the first $q$ columns of $\bm U$.

As a result of the compact SVD, we need calculate $\bm X_c \bm U_1$ only once
per cross-validation fold, requiring $O(p q N) \approx (p N^2)$ calculations.
Hence, the computational costs of expensive calculations, such as matrix
inverses and determinants, are greatly reduced because they are performed in the
$q$-dimensional subspace. Similarly, we reduce the dimension of the test data
set by calculating $\bm X_{test} \bm U_1$ once per fold. Conveniently, we see
that the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is
$\bm X_k \bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after
the initial calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm
Q_k$ requires $O(n_k q^2)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k \times
  n_k}$, both its determinant and inverse require $O(n_k^3)$
operations. Consequently, $\bm W_k^{-1}$ requires $O(n_k q^2)$ operations. Also,
the inverse and determinant of the diagonal matrix $\bm \Gamma_k \in
\mathbb{R}_{q \times q}$ requires $O(q)$ operations. Therefore, the $| \bm W_k
|$ requires $O(q n_k^3)$ operations.

The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
the selection of $\lambda$ and $\gamma$ via cross-validation because the most
time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
and $\gamma$ without repeating the costly computations.
